---
title: "modeling sampling"
author: "Linus"
date: "4 5 2021"
output: html_document
---

# load packages

```{r}
pkgs <- c("tidyverse")
lapply(pkgs, library, character.only = T)
```


# generate a lottery 

A lottery is defined by a) a vector of possible outcomes (Win, No Win), b) an outcome distribution (Probability of Win vs. No win), and c) the number of elements in the lottery (i.e., the maximum amount of draws from that lottery without replacement)

```{r}
set.seed(34)
outcomes <- c("w", "~w") # outcomes
p <- .20 # probability / outcome distribution (for 2 outcomes where p(~w) = 1-p(w))  
N <- 100 # number of elements in the lottery

# generate lottery
lot <- rep(outcomes, 
           times = c(p*N, (1-p)*N)
           ) 
lot <- factor(sample(lot)) # randomize order of elements
lot %>% table #  check underlying outcome distribution
```

# simulate sampling from a lottery 

Sampling is defined by a more or less complex sampling process, which produces single samples or subsets of the elements within a lotteries and then integrates them into a composite. A trivial sampling strategy is to draw a specified number of samples from a specified lottery with no further dependencies: 

## summary strategy with integration after termination

```{r}

# Sampling process 
set.seed(5) 
n <- 10 # number of single samples to be drawn from lottery
smpl <- sample(x = lot, # sample from lottery 
                size = n, 
                replace = FALSE # drawn outcomes elements are not replaced
                )
smpl
smpl %>% table

# Preference building 
C <- sum(recode(smpl, "~w" = 0, "w" = 1))  
C # composite score as number of experienced wins (expected value)
C/n # ... as experienced winning probability
```


## summary strategy with sequential sample integration (mean updating)

```{r}
lot <- seq(1,100,1) # lottery of 100 outcome elements 
n <- rep(1, 10) # sample size = 10
smpl <- vector("numeric", length(n)) # vector for unique samples
mean <- vector("numeric", length(n)) # vector for sample means

set.seed(7)

for (i in seq_along(n)) { # loop/sample n = 10 times
  
  smpl[i] <- sample(lot, # sample from lottery
                    size = n[i] # sample one element a time
                    )  
  
  lot <- lot[-(match(smpl[i], lot))] # no replacement/remove from lottery: 1st element in lottery vector that matches the element that was just sampled 

mean[i] <- mean(smpl[1:i]) # calculate mean of the current sample
  
}

output <- data.frame(ind = seq(1,length(nsmall),1), smpl, mean) # summary table of sampling process
output

output_l <- gather(output, stat, value, smpl:mean, factor_key=TRUE) # wide to long

graph <- output_l %>%
  ggplot(.) +
  geom_hline(yintercept = mean(lot)) + # plot lottery mean
  geom_point(aes(x = ind, y = value, color = stat))+ # plot single samples and corresponding mean updates
  geom_line(aes(x = ind, y = value, color = stat)) 
graph
```



